---
title: "Annulus Formation Full Analysis Code"
author: "LGCarlson"
date: "7/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
```

```{r}
theme_set(theme_light())

#set personalized plotting theme
LGCtheme <- theme_light() + theme(panel.background = element_blank(), 
                panel.grid = element_blank(), 
                axis.text = element_text(size = 13), 
                axis.title = element_text(size = 14), 
                legend.text = element_text(size = 13), 
                legend.title = element_text(size = 14), 
                strip.background = element_blank(), 
                strip.text = element_text(size = 13, color = "black"))
```


```{r, echo=FALSE}
formationdf<-read_csv(here::here("annulus_data", "formation_df.csv"))
```

#### Formation data metadata

* JoinID - ID composed of smolt year (release year) and Carlin tag number

* Batch_TagID - ID composed of batch number and Carlin tag number

* marine.circ - number of marine circuli total

* FMC - distance between scale focus and first marine circulus (mm)

* Radius - distance between scale focus and radius (scale edge) (mm)

* SeaAge - age at time of recapture based on number of marine annuli

* M1.incr - distance between first marine circulus and first marine annulus (mm)

* M2.incr - distance between first marine annulus and second marine annulus (mm)

* M1.circ - number of circuli between first marine circulus and first marine annulus

* M2.circ - number of circuli between first marine annulus and second marine annulus

* DaysatSea - number of days between release date and recapture date

* ReleaseDate - date that tagged smolt was released into the river (yyyy-mm-dd)

* RecaptureDate- date that tagged smolt was recaptured (yyyy-mm-dd)

* annulustype - modified later to represent labels used in manuscript (homewater and at sea rather than sea age)

* numannuli - descriptive columns that essentially repeat SeaAge

* Type - river or distant recapture




Clean data to remove "impossible" events

```{r, echo=F, message = F, warning = F}
# a fish cannot be at sea for more than 730 days and only have formed one annulus - recapture or release data must be incorrect
mistakeA <- formationdf %>% 
  filter(SeaAge == 1 & DaysatSea > 730) %>% 
  mutate(mistake = "at sea more than 730 days with only 1 annulus") 

# a fish cannot be at sea for less than 730 days and have already formed two annuli 
mistakeB <- formationdf %>% 
  filter(SeaAge == 2 & DaysatSea < 730) %>% 
  mutate(mistake = "at sea less than 730 days and already has 2 annuli") 

# a fish cannot be at sea for more than 350 days and not yet formed an annulus
mistakeC<- formationdf %>% 
  filter(SeaAge == 0 & DaysatSea > 300) %>% 
  mutate(mistake = "at sea more than 350 days with no annulus") 

# dataframe of "incorrect" or "impossible" data
mistakeFA<-bind_rows(mistakeA, mistakeB, mistakeC)

spawning1SW<-formationdf %>% 
  filter(numannuli == 1 & Type == "river") 

sea2SW<-formationdf %>% 
  filter(numannuli == 2 & Type == "distant") 

# overwrite formation data after removing mistake data and getting rid of fish caught as postsmolts
formationdf<-formationdf %>% 
  anti_join(mistakeFA, by = "JoinID") %>% 
  anti_join(spawning1SW, by = "JoinID") %>% 
  anti_join(sea2SW, by = "JoinID") %>% 
  filter(SeaAge > 0) %>% 
  mutate(annulustype = ifelse(Type == "river", "Homewater recaptures", "At sea recaptures"))
```


```{r}
table(formationdf$numannuli)
table(formationdf$annulustype)
length(unique(formationdf$JoinID))
```


The average deposition rate is 11.4, but deposition rate varies by number of marine annuli

```{r, echo=F, message = F, warning = F}
dayspercirc<-formationdf %>% 
  mutate(dayspercirc = DaysatSea/marine.circ) %>% 
  group_by(annulustype) %>% 
  dplyr::summarise(meancirday = mean(dayspercirc), sd = sd(dayspercirc), n = n(), 
                   `5%` = quantile(dayspercirc,probs=c(0.05)), `95%` = quantile(dayspercirc,probs=c(0.95)),
                   meandaysatsea = mean(DaysatSea), sddays = sd(DaysatSea)) 

dayspercirc
```

```{r, echo=F, message = F, warning = F, fig.height=2.75, fig.width=5, fig.align="center"}
formationdf %>% 
  mutate(dayspercirc = DaysatSea/marine.circ) %>% 
  left_join(dayspercirc, by = "annulustype") %>% 
  ggplot(aes(x=dayspercirc, fill=annulustype)) + 
  geom_histogram(color = "black", bins=30) + facet_wrap(~annulustype) +
  scale_fill_manual(values = c("#bdbdbd","#717171")) +
  geom_vline(aes(xintercept = meancirday), lty = 2, size = 0.7, color = "#2c2c2c") + 
  labs(x = "Deposition rate (days per circulus)", y = "n") +
  theme(panel.grid = element_blank()) + 
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 13)) +
  scale_x_continuous(breaks = c(6,9,12,15,18)) + LGCtheme +
  theme(legend.position = "none") 
  
```


### Calculate annulus formation date using individual average deposition rate

* last.annulus - M1 if at sea recapture or M2 if homewater recapture

* last.annulus.which - either "First marine annulus" or "Second marine annulus"

* after.growth - number of circuli after the last.annulus

* afterm1 - number of circuli after the first marine annulus

```{r, echo=F, message = F, warning = F}
aftergrowth_df<-formationdf %>% 
  mutate(last.annulus = ifelse(is.na(M2.incr) == FALSE, M2.circ,M1.circ)) %>% 
  mutate(last.annulus.which = ifelse(is.na(M2.incr) == FALSE, "Second marine annulus", "First marine annulus")) %>% 
  mutate(after.growth = marine.circ - last.annulus) %>% 
  mutate(afterm1 = marine.circ - M1.circ)
```


```{r}
anlabs<-as_labeller(c("M1DOY"="First marine annulus", "M2DOY"="Second marine annulus"))
```



## Individual constant scenario



```{r, echo = F, warning = F, fig.width=7.5, fig.height=3.75, fig.align="center"}
anndate<-aftergrowth_df %>% 
  mutate(dayspercirc = DaysatSea/marine.circ) %>% 
  mutate(dayssinceM1 = afterm1 * dayspercirc) %>% 
  mutate(dayssinceM2 = ifelse(last.annulus.which == "Second marine annulus", after.growth * dayspercirc, NA)) %>% 
  mutate(M1date = RecaptureDate - dayssinceM1) %>% 
  mutate(M2date = RecaptureDate - dayssinceM2) %>% 
  mutate(M1annmonth = lubridate::month(M1date)) %>% 
  mutate(M2annmonth = lubridate::month(M2date)) %>% 
  mutate(M1DOY = lubridate::yday(M1date)) %>% 
  mutate(M2DOY = lubridate::yday(M2date)) %>% 
  mutate(`First annulus` = ifelse(M1annmonth > 10, M1DOY-365, M1DOY)) %>% 
  mutate(`Second annulus` = ifelse(M2annmonth > 10, M2DOY-365, M2DOY)) %>% 
  gather(key = "Annulus", value = "DOY", 31:32) 

anndate %>%
  dplyr::group_by(annulustype, Annulus) %>% 
  summarise(mean = mean(DOY,na.rm=T), sd = sd(DOY,na.rm=T),
            `5%` = quantile(DOY,probs=c(0.05),na.rm=T), 
            `95%` = quantile(DOY,probs=c(0.95),na.rm=T)) %>% 
  dplyr::rename("Recpature Type" = "annulustype") %>% 
  drop_na()
```







