---
title: "Annulus Formation Full Analysis Code"
author: "LGCarlson"
date: "7/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
```

```{r}
theme_set(theme_light())

#set personalized plotting theme
LGCtheme <- theme_light() + theme(panel.background = element_blank(), 
                panel.grid = element_blank(), 
                axis.text = element_text(size = 13), 
                axis.title = element_text(size = 14), 
                legend.text = element_text(size = 13), 
                legend.title = element_text(size = 14), 
                strip.background = element_blank(), 
                strip.text = element_text(size = 13, color = "black"))
```


```{r, echo=FALSE}
Carlin_data<-read_csv(here::here("annulus_data", "formation_df.csv"))
```

#### Formation data metadata

* JoinID - ID composed of smolt year (release year) and Carlin tag number

* Batch_TagID - ID composed of batch number and Carlin tag number

* marine.circ - number of marine circuli total

* FMC - distance between scale focus and first marine circulus (mm)

* Radius - distance between scale focus and radius (scale edge) (mm)

* SeaAge - age at time of recapture based on number of marine annuli

* M1.incr - distance between first marine circulus and first marine annulus (mm)

* M2.incr - distance between first marine annulus and second marine annulus (mm)

* M1.circ - number of circuli between first marine circulus and first marine annulus

* M2.circ - number of circuli between first marine annulus and second marine annulus

* DaysatSea - number of days between release date and recapture date

* ReleaseDate - date that tagged smolt was released into the river (yyyy-mm-dd)

* RecaptureDate- date that tagged smolt was recaptured (yyyy-mm-dd)

* annulustype - modified later to represent labels used in manuscript (homewater and at sea rather than sea age)

* numannuli - descriptive columns that essentially repeat SeaAge

* Type - river or distant recapture




Clean data to remove "impossible" events

```{r, echo=F, message = F, warning = F}
# a fish cannot be at sea for more than 730 days and only have formed one annulus - recapture or release data must be incorrect
mistakeA <- Carlin_data %>% 
  filter(SeaAge == 1 & DaysatSea > 730) %>% 
  mutate(mistake = "at sea more than 730 days with only 1 annulus") 

# a fish cannot be at sea for less than 730 days and have already formed two annuli 
mistakeB <- Carlin_data %>% 
  filter(SeaAge == 2 & DaysatSea < 730) %>% 
  mutate(mistake = "at sea less than 730 days and already has 2 annuli") 

# a fish cannot be at sea for more than 350 days and not yet formed an annulus
mistakeC<- Carlin_data %>% 
  filter(SeaAge == 0 & DaysatSea > 300) %>% 
  mutate(mistake = "at sea more than 350 days with no annulus") 

# dataframe of "incorrect" or "impossible" data
mistakeFA<-bind_rows(mistakeA, mistakeB, mistakeC)

spawning1SW<-Carlin_data %>% 
  filter(numannuli == 1 & Type == "river") 

sea2SW<-Carlin_data %>% 
  filter(numannuli == 2 & Type == "distant") 

# overwrite formation data after removing mistake data and getting rid of fish caught as postsmolts
formationdf<-Carlin_data %>% 
  anti_join(mistakeFA, by = "JoinID") %>% 
  anti_join(spawning1SW, by = "JoinID") %>% 
  anti_join(sea2SW, by = "JoinID") %>% 
  filter(SeaAge > 0) %>% 
  mutate(annulustype = ifelse(Type == "river", "Homewater recaptures", "At sea recaptures"))
```


```{r}
table(formationdf$numannuli)
table(formationdf$annulustype)
length(unique(formationdf$JoinID))
```


The average deposition rate is 11.4, but deposition rate varies by number of marine annuli

```{r, echo=F, message = F, warning = F}
dayspercirc<-formationdf %>% 
  mutate(dayspercirc = DaysatSea/marine.circ) %>% 
  group_by(annulustype) %>% 
  dplyr::summarise(meancirday = mean(dayspercirc), sd = sd(dayspercirc), n = n(), 
                   `5%` = quantile(dayspercirc,probs=c(0.05)), `95%` = quantile(dayspercirc,probs=c(0.95)),`97%` = quantile(dayspercirc,probs=c(0.97)),
                   meandaysatsea = mean(DaysatSea), sddays = sd(DaysatSea)) 

dayspercirc
```

```{r, echo=F, message = F, warning = F, fig.height=2.75, fig.width=5, fig.align="center"}
formationdf %>% 
  mutate(dayspercirc = DaysatSea/marine.circ) %>% 
  left_join(dayspercirc, by = "annulustype") %>% 
  ggplot(aes(x=dayspercirc, fill=annulustype)) + 
  geom_histogram(color = "black", bins=30) + facet_wrap(~annulustype) +
  scale_fill_manual(values = c("#bdbdbd","#717171")) +
  geom_vline(aes(xintercept = meancirday), lty = 2, size = 0.7, color = "#2c2c2c") + 
  labs(x = "Deposition rate (days per circulus)", y = "n") +
  theme(panel.grid = element_blank()) + 
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 13)) +
  scale_x_continuous(breaks = c(6,9,12,15,18)) + LGCtheme +
  theme(legend.position = "none") +
  theme(panel.border = element_blank(), axis.line = element_line(color = "lightgrey"))


#ggsave(filename = "rate_hist.pdf", plot = last_plot(), height = 3, width = 5.25)  
```


Calculate annulus formation date using individual average deposition rate

New columns added:

* last.annulus - M1 if at sea recapture or M2 if homewater recapture

* last.annulus.which - either "First marine annulus" or "Second marine annulus"

* after.growth - number of circuli after the last.annulus

* afterm1 - number of circuli after the first marine annulus

```{r, echo=F, message = F, warning = F}
aftergrowth_df<-formationdf %>% 
  mutate(last.annulus = ifelse(is.na(M2.incr) == FALSE, M2.circ,M1.circ)) %>% 
  mutate(last.annulus.which = ifelse(is.na(M2.incr) == FALSE, "Second marine annulus", "First marine annulus")) %>% 
  mutate(after.growth = marine.circ - last.annulus) %>% 
  mutate(afterm1 = marine.circ - M1.circ)
```


```{r}
anlabs<-as_labeller(c("M1DOY"="First marine annulus", "M2DOY"="Second marine annulus"))
```



## Individual constant scenario



```{r, echo = F, warning = F, fig.width=7.5, fig.height=3.75, fig.align="center"}
anndate<-aftergrowth_df %>% 
  mutate(dayspercirc = DaysatSea/marine.circ) %>% 
  mutate(dayssinceM1 = afterm1 * dayspercirc) %>% 
  mutate(dayssinceM2 = ifelse(last.annulus.which == "Second marine annulus", after.growth * dayspercirc, NA)) %>% 
  mutate(M1date = RecaptureDate - dayssinceM1) %>% 
  mutate(M2date = RecaptureDate - dayssinceM2) %>% 
  mutate(M1annmonth = lubridate::month(M1date)) %>% 
  mutate(M2annmonth = lubridate::month(M2date)) %>% 
  mutate(M1DOY = lubridate::yday(M1date)) %>% 
  mutate(M2DOY = lubridate::yday(M2date)) %>% 
  mutate(`First annulus` = ifelse(M1annmonth > 10, M1DOY-365, M1DOY)) %>% 
  mutate(`Second annulus` = ifelse(M2annmonth > 10, M2DOY-365, M2DOY)) %>% 
  gather(key = "Annulus", value = "DOY", 31:32) 

anndate %>%
  dplyr::group_by(annulustype, Annulus) %>% 
  summarise(mean = mean(DOY,na.rm=T), sd = sd(DOY,na.rm=T),
            `5%` = quantile(DOY,probs=c(0.05),na.rm=T), 
            `95%` = quantile(DOY,probs=c(0.95),na.rm=T)) %>% 
  dplyr::rename("Recpature Type" = "annulustype") %>% 
  drop_na()

indcont_scenario<-anndate %>% 
  dplyr::select(JoinID, annulustype, Annulus, DOY) %>% 
  dplyr::rename("Individual constant" = "DOY")

```



## Seasonally variable scenario


Calculate intercirculus spacings from full scale (circulus measurement) data.

```{r, warning=F, message=F}
full_circ<-read_csv(here::here("annulus_data","Carlin_full_circuli_dataset.csv")) %>% 
  distinct(JoinID, .keep_all = T) %>% 
  dplyr::select(-X1) %>% 
  semi_join(formationdf, by = "JoinID")



spacinglist<-list()
IDvec<-as.vector(unique(full_circ$JoinID))

#calculate the differenced spacings
for(i in IDvec){
  
  temp_df<-full_circ %>% 
    distinct(JoinID, .keep_all = T) %>% 
    filter(JoinID == i)

# create a vector of circulus spacings
  vect<-as.numeric(temp_df[,2:127])
  vect<-vect[!is.na(vect)]

# create a vector of differenced spacings
  spacingraw<-diff(vect)
  
spacingdf<-tibble(JoinID = i, spacing = spacingraw) %>% 
    tibble::rowid_to_column("circ_num") %>% 
    dplyr::select(JoinID, circ_num, spacing)
  
spacinglist[[i]]<-spacingdf
}

spacing_df<-do.call(rbind,spacinglist) %>% 
  as_tibble() 
```


Apply seasonal variability to circulus deposition rate based on intercirculus spacing patterns.

```{r, warning=F, message=F}
ann<-read_csv(here::here("annulus_data","Carlin_growth_markers.csv")) %>% 
  distinct(JoinID, .keep_all = T) %>% 
  dplyr::select(-X1) %>% 
  semi_join(formationdf, by = "JoinID")


after_lastann<-list()
circrate_list<-list()

IDvec<-as.vector(unique(ann$JoinID))

for(i in IDvec){
  
# dataframe that contains only the circulus numbers of the FMC, M1, M2, and total number of circuli
  temp_annuli<-ann %>% 
    filter(JoinID == i) 

after_lastann[[i]]<- temp_annuli %>% 
  mutate(last_ann = ifelse(is.na(M2_c)==FALSE, M2_c, M1_c))  #add column of most recent annulus #

  
# add originally calculated data
  temp_rate<-anndate %>% 
    filter(JoinID == i) 
  
# keep only marine circuli
  temp_df<-spacing_df %>% 
    filter(JoinID == i) %>% 
    filter(circ_num >= temp_annuli$FMC_c)
  
# calculate deposition rate for each circulus based on spacings
  temp_df2<-temp_df %>% 
    mutate(centered = (spacing - mean(spacing))*-1) %>% 
    mutate(cent_rate = (centered*100)+temp_rate$dayspercirc) 


# sum of circ rates should = days at sea  
temp_rate$DaysatSea
sum(temp_df2$circ_rate)

circrate_list[[i]]<-temp_df2

}

seas_rate<-do.call(rbind,circrate_list) %>% 
   as_tibble()
seas_ann<-do.call(rbind,after_lastann) %>% 
   as_tibble()

```



Calculate new annulus formation dates based on these seasonally variable spacings.

```{r}
formationdates<-list()

for(i in IDvec){
  
# dataframe that contains only the circulus numbers of the FMC, M1, M2, total number of circuli, and circulus # of last annulus
l<-seas_ann %>% 
  distinct(JoinID, .keep_all = T) %>% 
  filter(JoinID == i) %>% 
  mutate(after_ann = total_c - last_ann) %>% 
  mutate(annulustype = ifelse(is.na(M2_c) == FALSE, "Homewater recaptures", "At sea recaptures"))

# filter out only "post-annulus" circuli
r<-seas_rate %>% 
  filter(JoinID == i) %>% 
  filter(circ_num %in% seq(l$last_ann, l$total_c,1))


s<-seas_rate %>% 
  filter(JoinID == i) %>% 
  filter(circ_num %in% seq(l$M1_c, l$total_c,1))

# add originally calculated data
temp_rate<-anndate %>% 
  distinct(JoinID, .keep_all = T) %>% 
  filter(JoinID == i)  


formationdates[[i]]<-tibble(JoinID= i, annulustype = l$annulustype, 
                            dayssinceM1 = sum(s$cent_rate), 
                            dayssinceM2 = sum(r$cent_rate), 
                            RecaptureDate = temp_rate$RecaptureDate, 
                            M1date = temp_rate$RecaptureDate - sum(s$cent_rate),
                            M2date = temp_rate$RecaptureDate - sum(r$cent_rate))
}

form_dates<-do.call(rbind,formationdates) %>% 
   as_tibble() %>% 
  mutate(M1annmonth = lubridate::month(M1date)) %>% 
  mutate(M2annmonth = lubridate::month(M2date)) %>% 
  mutate(M1DOY = lubridate::yday(M1date)) %>% 
  mutate(M2DOY = lubridate::yday(M2date)) %>% 
  mutate(`First annulus` = ifelse(M1annmonth > 10, M1DOY-365, M1DOY)) %>% 
  mutate(`Second annulus` = ifelse(M2annmonth > 10, M2DOY-365, M2DOY)) %>% 
  mutate(`Second annulus` = ifelse(annulustype == "At sea recaptures", NA, M2DOY)) %>% 
  gather(key = "Annulus", value = "DOY", 12:13) 


form_dates %>%
  group_by(annulustype, Annulus) %>% 
  summarise(mean = mean(DOY,na.rm=T), sd = sd(DOY,na.rm=T),
            `5%` = quantile(DOY,probs=c(0.05),na.rm=T), 
            `95%` = quantile(DOY,probs=c(0.95),na.rm=T)) %>%  
  dplyr::rename("Recpature Type" = "annulustype") %>% 
  drop_na()




seasvar_scenario<-form_dates %>% 
  dplyr::select(JoinID, annulustype, Annulus, DOY) %>% 
  dplyr::rename("Seasonally variable" = "DOY")

```


## Lifetime and seasonally variable scenario

```{r}
K_calc<-read_csv(here::here("annulus_data", "K_calc_data.csv"))

meanDAS<-K_calc %>% 
  group_by(SeaAge) %>% 
  summarise(meanDAS=mean(DaysatSea, na.rm = T)) %>% 
  mutate(meanDAS = round(meanDAS,0))

K_calc %>% 
  group_by(SeaAge) %>% 
  count()
```


```{r, fig.height=5, fig.width=4, fig.align="center"}

growth_mod <- glm(K_calc$marine.incr ~ poly(K_calc$DaysatSea,2))
summary(growth_mod)

pred_vals<-predict.glm(growth_mod,data.frame(x=K_calc$DaysatSea),type="response",se.fit = T)
critval <- 1.96 ## approx 95% CI
upr <- pred_vals$fit + (critval * pred_vals$se.fit)
lwr <- pred_vals$fit - (critval * pred_vals$se.fit)

pred<-tibble(DaysatSea = K_calc$DaysatSea, fit = pred_vals$fit, upr = upr, lwr = lwr) %>% 
  distinct(DaysatSea, .keep_all = T) %>% 
  arrange(DaysatSea)

adj2SW<-unique(pred$fit[pred$DaysatSea == meanDAS$meanDAS[3]])/unique(pred$fit[pred$DaysatSea == meanDAS$meanDAS[2]])
adj1SW<-adj2SW*0.9
inv1SW<-2-adj1SW


ggplot(K_calc, aes(DaysatSea, marine.incr)) +
  geom_point(size=1.6, color = "#C7C9CB") + 
  geom_line(data=pred, aes(x=DaysatSea, y=fit),size=0.7) + 
  geom_line(data=pred, aes(x=DaysatSea, y=upr),size=0.6, lty=2) + 
  geom_line(data=pred, aes(x=DaysatSea, y=lwr),size=0.6, lty=2) + LGCtheme +
  labs(x= "Days at Sea", y = "Marine growth increment", fill = "") +
  scale_x_continuous(breaks = c(0,250,500,750))+
  theme(panel.border = element_blank(), axis.line = element_line(color = "lightgrey"))


#ggsave(filename = "asympt_growth.pdf", plot = last_plot(), height = 5, width = 4)  
```



```{r}
after_lastann<-list()
circrate_list<-list()

IDvec<-as.vector(unique(ann$JoinID))

for(i in IDvec){
  
# dataframe that contains only the circulus numbers of the FMC, M1, M2, and total number of circuli
  temp_annuli<-ann %>% 
    filter(JoinID == i) 

after_lastann[[i]]<- temp_annuli %>% 
  mutate(last_ann = ifelse(is.na(M2_c)==FALSE, M2_c, M1_c))   #add column of most recent annulus #

  
# add originally calculated data to get average dep rate
  temp_rate<-anndate %>% 
    filter(JoinID == i) 

# age adjusted rate happens here
  temp_df<-spacing_df %>% 
    filter(JoinID == i) %>% 
    mutate(SeaAge = unique(temp_rate$SeaAge)) %>% 
    filter(circ_num >= temp_annuli$FMC_c)    # keep only marine circuli

  
if(is.na(temp_annuli$M2_c) == F){
  temp_df1<-temp_df %>% 
    mutate(which_circ = ifelse(circ_num >= temp_annuli$M2_c, "after M2", NA)) %>% 
    mutate(which_circ = ifelse(circ_num < temp_annuli$M2_c, "M1-M2", which_circ)) %>% 
    mutate(which_circ = ifelse(circ_num <= temp_annuli$M1_c, "before M1", which_circ)) %>% 
    mutate(which_rate = ifelse(circ_num >= temp_annuli$M2_c, (adj2SW*temp_rate$dayspercirc), NA)) %>% 
    mutate(which_rate = ifelse(circ_num <= temp_annuli$M1_c, ((inv1SW*temp_rate$dayspercirc)), which_rate)) 

  temp_df2<- temp_df1 %>% 
    mutate(daysused = (unique(temp_rate$DaysatSea)-sum(temp_df1$which_rate, 
                                               na.rm = T))/sum(temp_df1$which_circ == "M1-M2")) %>% 
    mutate(which_rate = ifelse(which_circ == "M1-M2", daysused, which_rate)) %>% 
    dplyr::select(-daysused)

}
  
  else{
  temp_df1<-temp_df %>% 
    mutate(which_circ = ifelse(circ_num <= temp_annuli$M1_c, "before M1", "after M1")) %>% 
    mutate(which_rate = ifelse(circ_num <= temp_annuli$M1_c, NA,
                               (adj1SW*temp_rate$dayspercirc))) 

  temp_df2<- temp_df1 %>%  
    mutate(daysused = (unique(temp_rate$DaysatSea)-sum(temp_df1$which_rate, 
                        na.rm = T))/(unique(temp_rate$marine.circ)-sum(temp_df1$which_circ == "after M1"))) %>% 
    mutate(which_rate = ifelse(which_circ == "before M1", daysused, which_rate)) %>% 
    dplyr::select(-daysused)
}

# calculate deposition rate for each circulus based on spacings
  temp_df3<-temp_df2 %>% 
    mutate(centered = (spacing - mean(spacing))*-1) %>% 
    mutate(cent_rate = (centered*100)+which_rate)


# sum of circ rates should = days at sea, some rounding error with rate at age adjustment
temp_rate$DaysatSea
sum(temp_df3$cent_rate)

circrate_list[[i]]<-temp_df3 %>% 
  mutate(diff=(unique(temp_rate$DaysatSea))-sum(temp_df3$cent_rate)) %>% 
  mutate(diff = round(diff,0))

}

life_rate<-do.call(rbind,circrate_list) %>% 
   as_tibble()
life_ann<-do.call(rbind,after_lastann) %>% 
   as_tibble()

```




```{r}
formationdates<-list()

for(i in IDvec){
  
# dataframe that contains only the circulus numbers of the FMC, M1, M2, total number of circuli, and circulus # of last annulus
l<-life_ann %>% 
  distinct(JoinID, .keep_all = T) %>% 
  filter(JoinID == i) %>% 
  mutate(after_ann = total_c - last_ann) %>% 
  mutate(annulustype = ifelse(is.na(M2_c) == FALSE, "Homewater recaptures", "At sea recaptures")) %>% 
  mutate(firstann = ifelse(is.na(M2_c) == FALSE, M2_c-M1_c, NA))

# filter out only "post-annulus" circuli
r<-life_rate %>% 
  filter(JoinID == i) %>% 
  filter(circ_num %in% seq(l$last_ann, l$total_c,1))


s<-life_rate %>% 
  filter(JoinID == i) %>% 
  filter(circ_num %in% seq(l$M1_c, l$total_c,1))

# add originally calculated data
temp_rate<-anndate %>% 
  distinct(JoinID, .keep_all = T) %>% 
  filter(JoinID == i)  

formationdates[[i]]<-tibble(JoinID= i, annulustype = l$annulustype, 
                            dayssinceM1 = sum(s$cent_rate), 
                            dayssinceM2 = sum(r$cent_rate), 
                            RecaptureDate = temp_rate$RecaptureDate, 
                            M1date = temp_rate$RecaptureDate - sum(s$cent_rate),
                            M2date = temp_rate$RecaptureDate - sum(r$cent_rate))
}

form_dates<-do.call(rbind,formationdates) %>% 
   as_tibble() %>% 
  mutate(M1annmonth = lubridate::month(M1date)) %>% 
  mutate(M2annmonth = lubridate::month(M2date)) %>% 
  mutate(M1DOY = lubridate::yday(M1date)) %>% 
  mutate(M2DOY = lubridate::yday(M2date)) %>% 
  mutate(`First annulus` = ifelse(M1annmonth > 10, M1DOY-365, M1DOY)) %>% 
  mutate(`Second annulus` = ifelse(M2annmonth > 10, M2DOY-365, M2DOY)) %>% 
  mutate(`Second annulus` = ifelse(annulustype == "At sea recaptures", NA, M2DOY)) %>% 
  gather(key = "Annulus", value = "DOY", 12:13) 


form_dates %>%
  group_by(annulustype, Annulus) %>% 
  summarise(mean = mean(DOY,na.rm=T), sd = sd(DOY,na.rm=T),
            `5%` = quantile(DOY,probs=c(0.05),na.rm=T), 
            `95%` = quantile(DOY,probs=c(0.95),na.rm=T)) %>%  
  dplyr::rename("Recpature Type" = "annulustype") %>% 
  drop_na()



seaslife_scenario<-form_dates %>% 
  dplyr::select(JoinID, annulustype, Annulus, DOY) %>% 
  dplyr::rename("Seasonal and lifetime variable" = "DOY")

```





## Winter constant scenario (14 days/circulus)

```{r, echo = F, warning = F, fig.width=7.5, fig.height=3.75, fig.align="center"}
anndate_14<-aftergrowth_df %>% 
  mutate(dayspercirc = 14) %>% 
  mutate(dayssinceannulus = after.growth * dayspercirc) %>% 
  mutate(annulusdate = RecaptureDate - dayssinceannulus) %>% 
  mutate(M1annmonth = ifelse(annulustype == "At sea recaptures", lubridate::month(annulusdate), NA)) %>% 
  mutate(M2annmonth = ifelse(annulustype == "Homewater recaptures", lubridate::month(annulusdate), NA)) %>% 
  mutate(M1DOY = ifelse(annulustype == "At sea recaptures",lubridate::yday(annulusdate),NA)) %>% 
  mutate(M2DOY = ifelse(annulustype == "Homewater recaptures",lubridate::yday(annulusdate), NA)) %>% 
  mutate(`First annulus` = ifelse(M1annmonth > 10, M1DOY-365, M1DOY)) %>% 
  mutate(`Second annulus` = ifelse(M2annmonth > 10, M2DOY-365, M2DOY)) %>% 
  gather(key = "Annulus", value = "DOY", 29:30) 

anndate_14 %>%
  group_by(annulustype, Annulus) %>% 
  summarise(mean = mean(DOY,na.rm=T), sd = sd(DOY,na.rm=T),
            `5%` = quantile(DOY,probs=c(0.05),na.rm=T), 
            `95%` = quantile(DOY,probs=c(0.95),na.rm=T)) %>%  
  dplyr::rename("Recpature Type" = "annulustype") %>% 
  drop_na()

wintercont_scenario<-anndate_14 %>% 
  dplyr::select(JoinID, annulustype, Annulus, DOY) %>% 
  dplyr::rename("Constant (14 days/circ)" = "DOY") %>% 
  drop_na()
```



## Summer constant scenario (7 days/circulus)

```{r, echo = F, warning = F}
anndate_7<-aftergrowth_df %>% 
  mutate(dayspercirc = 7) %>% 
  mutate(dayssinceannulus = after.growth * dayspercirc) %>% 
  mutate(annulusdate = RecaptureDate - dayssinceannulus) %>% 
  mutate(M1annmonth = ifelse(annulustype == "At sea recaptures", lubridate::month(annulusdate), NA)) %>% 
  mutate(M2annmonth = ifelse(annulustype == "Homewater recaptures", lubridate::month(annulusdate), NA)) %>% 
  mutate(M1DOY = ifelse(annulustype == "At sea recaptures",lubridate::yday(annulusdate),NA)) %>% 
  mutate(M2DOY = ifelse(annulustype == "Homewater recaptures",lubridate::yday(annulusdate), NA)) %>% 
  mutate(`First annulus` = ifelse(M1annmonth > 10, M1DOY-365, M1DOY)) %>% 
  mutate(`Second annulus` = ifelse(M2annmonth > 10, M2DOY-365, M2DOY)) %>% 
  gather(key = "Annulus", value = "DOY", 29:30) 

anndate_7 %>%
  group_by(annulustype, Annulus) %>% 
  summarise(mean = mean(DOY,na.rm=T), sd = sd(DOY,na.rm=T),
            `5%` = quantile(DOY,probs=c(0.05),na.rm=T), 
            `95%` = quantile(DOY,probs=c(0.95),na.rm=T)) %>%  
  dplyr::rename("Recpature Type" = "annulustype") %>% 
  drop_na()


summercont_scenario<-anndate_7 %>% 
  dplyr::select(JoinID, annulustype, Annulus, DOY) %>% 
  dplyr::rename("Constant (7 days/circ)" = "DOY") %>% 
  drop_na()
```



```{r,fig.height=3, fig.width=14}
full_formation_data<-indcont_scenario %>%  
  left_join(seasvar_scenario, by=c("JoinID", "Annulus", "annulustype")) %>% 
  left_join(seaslife_scenario, by=c("JoinID", "Annulus", "annulustype")) %>% 
  left_join(wintercont_scenario, by=c("JoinID", "Annulus", "annulustype")) %>% 
  left_join(summercont_scenario, by=c("JoinID", "Annulus", "annulustype")) %>% 
  mutate(whichann = paste(annulustype, Annulus, sep = " - ")) %>% 
  gather(key="scenario", value="DOY", 4:8) 



mylabels<-c("Homewater recaptures \n Second marine annulus","Homewater recaptures \n First marine annulus","At sea recaptures \n First marine annulus")

full_formation_data %>% 
  drop_na() %>% 
  mutate(whichann = fct_relevel(whichann, c("Homewater recaptures - Second annulus","Homewater recaptures - First annulus",
                                      "At sea recaptures - First annulus"))) %>%
  mutate(scenario = fct_relevel(scenario, c("Individual constant", "Seasonally variable", "Seasonal and lifetime variable", 
                                            "Constant (14 days/circ)", "Constant (7 days/circ)"))) %>% 
  ggplot(aes(y=whichann, x=DOY, fill=as.factor(annulustype))) + 
  ggridges::stat_density_ridges(quantile_lines = TRUE, quantiles = c(0.05, 0.5, 0.95), 
                                scale=0.81, alpha = 0.8, rel_min_height = 0.01) +
  geom_vline(xintercept = -12, lty=2, color="#2c2c2c") +
  annotate("text",x=-31, y="Homewater recaptures - First annulus", label = "Winter solstice", angle=90, color = "#2c2c2c") +
  facet_wrap(~scenario, ncol=5) +
  scale_fill_manual(values = c("#bdbdbd","#717171")) +
  scale_x_continuous(limits = c(-80, 280), breaks = c(-30,60, 152,244), 
                     labels = c("Dec 1","Mar 1","Jun 1","Sep 1")) + 
  scale_y_discrete(labels = mylabels) + LGCtheme +
  labs(x = "Predicted date of annulus completion", y = "") +
  theme(legend.position = "none")   + 
  theme(panel.border = element_blank(), axis.line = element_line(color = "lightgrey"))



###ggsave(filename = "completion_ridges.pdf", plot = last_plot(), height = 4, width = 18)
```


