---
title: "Untitled"
author: "LGCarlson"
date: "7/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
```


```{r}
formation_df<-read_csv(here::here("annulus_data", "formation_df.csv"))
```

#### Formation data metadata

* JoinID - ID composed of smolt year (release year) and Carlin tag number

* Batch_TagID - ID composed of batch number and Carlin tag number

* marine.circ - number of marine circuli total

* FMC - distance between scale focus and first marine circulus (mm)

* Radius - distance between scale focus and radius (scale edge) (mm)

* SeaAge - age at time of recapture based on number of marine annuli

* M1.incr - distance between first marine circulus and first marine annulus (mm)

* M2.incr - distance between first marine annulus and second marine annulus (mm)

* M1.circ - number of circuli between first marine circulus and first marine annulus

* M2.circ - number of circuli between first marine annulus and second marine annulus

* DaysatSea - number of days between release date and recapture date

* ReleaseDate - date that tagged smolt was released into the river (yyyy-mm-dd)

* RecaptureDate- date that tagged smolt was recaptured (yyyy-mm-dd)

* annulustype and numannuli - descriptive columns that essentially repeat SeaAge

* Type - river (homewater) or distant (at sea) recapture


```{r}

```

